%\documentclass[12pt]{scrreprt}
\documentclass[12pt]{report}

% language may be romanian or english (default is english)
% type may be bachelor or master (default is bachelor)
\usepackage[language=english, type=bachelor]{style}

%\geometry{a4paper,top=2.5cm,left=3cm,right=2.5cm,bottom=2.5cm}
%in style
%controlling the appearance of your headers and footers
\usepackage{fancyhdr}
\usepackage{caption}
\usepackage{subcaption}
\pagestyle{fancy}
\lhead{}
\chead{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.2pt}

\begin{document}

\specialization{COMPUTER SCIENCE}
\title{Integration of Neural Networks and Minimax Strategies in Chess Engine Development}
\author{Cadar Eduard}
\supervisor{Asist. univ. dr. Florentin Bota}
\maketitle

\begin{titlepage}
\begin{center}
    {\Large\textbf{
    UNIVERSITATEA BABE\c{S}-BOLYAI CLUJ-NAPOCA \\
    FACULTATEA DE MATEMATIC\u{A} \c{S}I INFORMATIC\u{A} \\
    SPECIALIZAREA INFORMATIC\u{A} \\
    }}
    \end{center}
    
    \vspace{10em}
    
    \begin{center}
    {\Huge\textbf{
    LUCRARE DE LICEN\c T\u A
    }}
    \end{center}
    
    \vspace{3em}
    
    \begin{center}
    {\huge\textbf{Integrarea Re\c{t}elelor Neuronale \c{s}i a Strategiilor Minimax \^{i}n Dezvoltarea unui Motor de \c{S}ah}}
    \end{center}
    
    \vspace{10em}
    
    \begin{flushleft}
    {\Large\textbf{
    Conduc\u{a}tor \c{s}tiin\c{t}ific \\ Asist. univ. dr. Florentin Bota
    }}
    \end{flushleft}
    
    \vspace{3em}
    
    \begin{flushright}
    {\Large\textit{
    Absolvent\\ Cadar Eduard
    }}
    \end{flushright}
    
    \vfill
    
    \begin{center}
    {\Large{\the\year}}
\end{center}
\end{titlepage}

\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\pagenumbering{roman} 

\cleardoublepage
ABSTRACT
\vspace{0.5cm}	
\hrule
\vspace{0.5cm}	
%\cleardoublepage

The usual way of searching for a move from a chess position is through tactics or strategy. Tactics are series of moves that would bring an immediate advantage, while strategy refers to a general 'sense' of advantage on the board, without too much calculation (having better developed pieces, better pawn structure etc.) \cite{klein2022neural}.

Chess engines are usually built using a minimax algorithm, along with an evaluation function that takes into consideration the placement of the pieces on the board. This approach is able to find the moves that would bring material gain (capturing pieces) in a given depth, but cannot find moves that would slowly improve the position. Another approach would be training a neural network on professional chess games, which would be able to find good positional moves, but may be weak at finding tactical moves.

An algorithm combining these two approaches should yield better results than each of them on their own.

\tableofcontents
\newpage
\listoffigures
\pagenumbering{arabic}

\input{chapters/chapter1_introduction}
%\addcontentsline{toc}{chapter}{Introducere}
%\addcontentsline{toc}{chapter}{Introduction}

\input{chapters/chapter2}
\input{chapters/chapter3}
\input{chapters/chapter4}
\input{chapters/chapter5}
\input{chapters/chapter6}
\input{chapters/chapter7_conclusions}
%\addcontentsline{toc}{chapter}{Concluzii}
%\addcontentsline{toc}{chapter}{Conclusions}

\bibliography{references}

\end{document}
